---
title: "scarch"
author: "Yuyang Sun"
date: "2023-11-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Library
```{r, echo= FALSE}
library(ggplot2)
library(lme4)
library(dplyr)
library(caret)
library(corrplot)
library(pROC)
library(viridis)
library(gplots)
```

## Load Data
### Add a year column for Multilevel Model
```{r}
# Load Data
bank <- read.csv("bank-full.csv" , sep = ";", header = TRUE)

# Initialize the year column
bank$year <- 2008

# Loop through the rows and increment the year after each December
for (i in 2:nrow(bank)) {
  if (tolower(bank$month[i]) == "jan" && tolower(bank$month[i-1]) == "dec") {
    bank$year[i] <- bank$year[i-1] + 1
  } else {
    bank$year[i] <- bank$year[i-1]
  }
}

# Check the result
head(bank)
table(bank$year)
```

### Adjust the format of variables
```{r, echo=FALSE}
# Convert binary variables to 0 and 1
bank$default <- ifelse(bank$default == "yes", 1, 0)
bank$housing <- ifelse(bank$housing == "yes", 1, 0)
bank$loan <- ifelse(bank$loan == "yes", 1, 0)
bank$y <- ifelse(bank$y == "yes", 1, 0)

# Convert categorical variables to dummy variables
bank <- bank %>% 
  mutate(across(.cols = c(job, marital, education, contact, poutcome), .fns = as.factor)) %>%
  mutate(across(.cols = c(day, month), .fns = as.factor)) %>%
  mutate(across(.cols = c(age, balance, duration, campaign, pdays, previous), .fns = as.numeric)) %>%
  mutate(day = as.numeric(as.factor(day)), month = as.numeric(as.factor(month)))

head(bank)
```

## Initital Setup

### Check NA's

```{r, echo=FALSE}
library(ggplot2)
library(reshape2)

# Melting the data
bank_melted <- melt(is.na(bank), id.vars = NULL)

# Creating a heatmap of NA values
ggplot(bank_melted, aes(Var2, Var1)) +
  geom_raster(aes(fill = value)) +
  scale_fill_manual(values = c("TRUE" = "red", "FALSE" = "grey")) +
  theme_minimal() +
  labs(x = "Variables", y = "Observations", title = "Heatmap of Missing Values") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```


```{r, echo=FALSE}
# Number of NA values for each variable
library(ggplot2)
library(dplyr)

# Bin the age data
bank_binned <- bank %>%
  mutate(age_bin = cut(age, breaks = seq(min(age), max(age), by = 5), include.lowest = TRUE)) %>%
  group_by(age_bin) %>%
  summarize(average_balance = mean(balance, na.rm = TRUE))

# Create the bar plot
ggplot(bank_binned, aes(x = age_bin, y = average_balance)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  theme_minimal() +
  labs(title = "Average Balance Grouped by Age Bins",
       x = "Age Bins",
       y = "Average Balance") +
  theme(plot.title = element_text(hjust = 0.5)) # Center the title

```

#### Cut the age into subgroups

```{r, echo=FALSE}

```

### Balance vs. Duration

```{r,warning=FALSE, echo=FALSE}
library(ggplot2)
library(dplyr)

# Bin the duration data and calculate the average balance for each bin
bank_binned <- bank %>%
  mutate(duration_bin = cut(duration, breaks = seq(0, max(duration, na.rm = TRUE), by = 180), include.lowest = TRUE)) %>%
  group_by(duration_bin) %>%
  summarize(average_balance = mean(balance, na.rm = TRUE))

# Bar plot of the average balance for each duration bin
ggplot(bank_binned, aes(x = duration_bin, y = average_balance)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  theme_minimal() +
  labs(title = "Average Balance Grouped by Duration Bins",
       x = "Duration Bins",
       y = "Average Balance") +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 45, vjust = 0.5))

```

### Duration vs. Loan, Housing
```{r, echo=FALSE}
ggplot(bank, aes(x = as.factor(loan), y = duration)) +
  geom_violin(trim = FALSE) +
  labs(title = "Violin Plot of Duration by Loan Status",
       x = "Loan Status (0 = No, 1 = Yes)",
       y = "Duration")

mean_duration_loan <- aggregate(duration ~ loan, data = bank, mean)
median_duration_loan <- aggregate(duration ~ loan, data = bank, median)
combined_duration_loan <- merge(mean_duration_loan, median_duration_loan, by = "loan")
names(combined_duration_loan) <- c("Loan", "Mean Duration", "Median Duration")
print(combined_duration_loan)

ggplot(bank, aes(x = as.factor(housing), y = duration)) +
  geom_violin(trim = FALSE) +
  labs(title = "Violin Plot of Duration by Housing Status",
       x = "Loan Status (0 = No, 1 = Yes)",
       y = "Duration")

mean_duration_housing <- aggregate(duration ~ housing, data = bank, mean)
median_duration_housing <- aggregate(duration ~ housing, data = bank, median)
combined_duration_housing <- merge(mean_duration_housing, median_duration_housing, by = "housing")
names(combined_duration_housing) <- c("housing", "Mean Duration", "Median Duration")
print(combined_duration_housing)
```

### Balance vs. Loan, Housing
```{r, echo=FALSE}
ggplot(bank, aes(x = as.factor(loan), y = balance)) +
  geom_violin(trim = FALSE) +
  labs(title = "Violin Plot of Balance by Loan Status",
       x = "Loan Status (0 = No, 1 = Yes)",
       y = "Balance")

mean_balance_loan <- aggregate(balance ~ loan, data = bank, mean)
median_balance_loan <- aggregate(balance ~ loan, data = bank, median)
combined_balance_loan <- merge(mean_balance_loan, median_balance_loan, by = "loan")
names(combined_balance_loan) <- c("Loan", "Mean Balance", "Median Balance")
print(combined_balance_loan)

ggplot(bank, aes(x = as.factor(housing), y = balance)) +
  geom_violin(trim = FALSE) +
  labs(title = "Violin Plot of Balance by Housing Status",
       x = "Loan Status (0 = No, 1 = Yes)",
       y = "Balance")

mean_balance_housing <- aggregate(balance ~ housing, data = bank, mean)
median_balance_housing <- aggregate(balance ~ housing, data = bank, median)
combined_balance_housing <- merge(mean_balance_housing, median_balance_housing, by = "housing")
names(combined_balance_housing) <- c("housing", "Mean Balance", "Median Balance")
print(combined_balance_housing)
```

### Correlation of Numeric Data
```{r warning=FALSE}
# Assuming you have already calculated the correlation matrix 'corr_matrix'
numerical_data <- bank[sapply(bank, is.numeric)]
corr_matrix <- cor(numerical_data, use = "complete.obs")

# Set the size of the plotting device (adjust width and height as needed)
options(repr.plot.width = 8, repr.plot.height = 8)

# Define a custom color palette (for example, using the "cividis" color palette)
custom_palette <- cividis(256, alpha = 1, begin = 0, end = 1, direction = 1)

# Create a smaller heatmap with adjusted settings and no dendrograms
heatmap.2(corr_matrix,
          col = custom_palette,
          main = "Correlation Heatmap",  # Remove the main title
          symm = TRUE,
          trace = "none",  # Hide row and column labels
          key = TRUE,      # Display a color legend
          key.title = "Frequency",
          key.title.cex = 0.2,  # Adjust the title font size
          key.subtitle = "Color Scale",
          key.subtitle.cex = 0.3,  # Adjust the subtitle font size
          cex.legend = 0.5,        # Adjust the size of the color legend key
          cex.axis = 0.25,         # Adjust axis label font size
          cex.main = 0.4,          # Adjust main title font size
          cex.lab = 0.25,          # Adjust axis labels font size
          lhei = c(0.02, 0.05),    # Adjust smaller row and column height
          lwid = c(0.02, 0.05),    # Adjust smaller row and column width
          cellnote = round(corr_matrix, 2),  # Display correlation values
          notecol = "white",       # Set the color of the correlation values
          notecex = 0.8,          # Adjust the correlation value font size
          symbreaks = FALSE,
          dendrogram = "none",
          )

```

### Check pday, poutcome
```{r}
summary(bank$pdays)
sum(is.na(bank$pdays))
sum(bank$pdays == -1)  # or any other special code

summary(bank$campaign)
table(bank$poutcome)

```

```{r}
model_pdays <- glm(y ~ pdays, data = bank, family = binomial)
model_campaign <- glm(y ~ campaign, data = bank, family = binomial)
model_poutcome <- glm(y ~ poutcome, data = bank, family = binomial)
summary(model_pdays)
summary(model_campaign)
summary(model_poutcome)
```
### Multicollinearity Check

```{r}
# Assuming you have a model built, e.g., logistic regression
model <- glm(y ~ ., data = bank, family = binomial)

# Calculate VIF
library(car)
vif(model)
# Values of VIF > 5 or 10 indicate high multicollinearity

```
### Interaction between job and education

```{r}
# Interaction plot for Job and Education
ggplot(bank, aes(x = job, fill = education)) +
  geom_bar(position = "dodge") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Interaction between Job and Education")

```


## Logistic Regression

```{r}
# Splitting data into training and test sets
set.seed(55)
training_rows <- createDataPartition(bank$y, p = 0.7, list = FALSE)
train_data <- bank[training_rows, ]
test_data <- bank[-training_rows, ]

# Build logistic regression model
model <- glm(y ~ ., data = train_data, family = binomial)
summary(model)

# Make predictions on the test set
predictions <- predict(model, newdata = test_data, type = "response")

# Convert predictions to binary format
predicted_class <- ifelse(predictions > 0.5, 1, 0)

# Confusion matrix and accuracy
confusionMatrix <- confusionMatrix(factor(predicted_class), factor(test_data$y))
print(confusionMatrix)

# ROC and AUC

roc_result <- roc(test_data$y, predictions)
auc(roc_result)
plot(roc_result)

# Build logistic regression model with two categorical predictors
model_categorical <- glm(y ~ default + job, data = train_data, family = binomial)

# Summary of the model
summary(model_categorical)


```

## Multilevel regression
```{r}
# Convert categorical variables to factors if they are not already
bank$education <- as.factor(bank$education)
bank$marital <- as.factor(bank$marital)
bank$loan <- as.factor(bank$loan)
bank$housing <- as.factor(bank$housing)

# Build the model
# Replace 'cluster_var' with your actual clustering variable
model <- glmer(y ~ age + education + marital + balance + loan + housing + (1 | year), 
               data = bank, 
               family = binomial)

# View the model summary
summary(model)
```


```{r}

```
